---
title: "WFS Multibeam Metadata Sheet"
author: "Alexander Ilich"
date: "`r format(Sys.time(), '%B %d, %Y %H:%M')`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE) #Default chunk options
```

This compilation of West Florida Shelf Multibeam data is largely for internal usage of the [C-SCAMP](https://www.marine.usf.edu/scamp/) and [COMIT](https://www.marine.usf.edu/comit/) groups at the University of South Florida College of Marine Science. This is a living document that will change as more data is collected, discovered, and reprocessed. The metadata csv/xlsx miles and the GIS vector files (shp and gpkg files) may be useful for others to understand the nature and location of various data sets that exist. The Server_Path represents the location of the data on the C-SCAMP/COMIT servers and are therefore not directly accessible to those outside the organization. In the long term, these data sets collected by C-SCAMP and COMIT will be made publicly available via [NCEI](https://www.ncei.noaa.gov/maps/bathymetry/). In the mean time, these data sets can be requested by emailing comit-data@usf.edu. Data sets collected by external groups may be available by at the specified URL in the metadata spreadsheet.


```{r include=FALSE}
library(tidyverse)
library(writexl)
library(terra)
library(sf)

mb_df<- read_csv("input_filepaths_sorted.csv")

if(file.exists("Multibeam_metadata.csv")){
  mb_old<- read_csv("Multibeam_metadata.csv") %>% select(ID, Server_Location)
} else{
  mb_old<- tibble(ID=NA_real_, Server_Location=NA_character_ , .rows = 0)
}
names(mb_old)[2]<- "Old_Server_Location"

mb_df<- bind_cols(mb_df[,1:which(names(mb_df)=="Type")], 
          tibble(Resolution=NA_real_,	Units= NA_character_,	Projection_Description=NA_character_,	Projection=NA_character_,	Ellipsoid=NA_character_, .rows = nrow(mb_df)),
          mb_df[,(which(names(mb_df)=="Type")+1):ncol(mb_df)]) #Add columns to fill in with new information
          
mb_df<- mb_df %>% left_join(mb_old, by="ID")

mb_df<- st_sf(cbind(mb_df, st_sf(geometry = st_sfc(lapply(1:nrow(mb_df), function(x) st_multipolygon()))))) #Make spatial
st_crs(mb_df)<- "EPSG:4326" #Add projection

if(file.exists("Multibeam_metadata.gpkg")){
  mb_shp<- st_read("Multibeam_metadata.gpkg")
  for (i in 1:nrow(mb_df)) {
    if(isTRUE(mb_df$Server_Location[i]==mb_df$Old_Server_Location[i])){
      curr_geom<- st_geometry(mb_shp[which(mb_shp$Server_Location == mb_df$Server_Location[i]),])
      if(!st_is_empty(curr_geom)){
      mb_df$geometry[i]<- curr_geom
      }
      }
    }
} # Use existing footprints if file has already been created

for (i in 1:nrow(mb_df)) {
  if(!is.na(mb_df$Server_Location[i])){
    if(str_detect(mb_df$Server_Location[i], "^D:")){
      prefix<-""} else{
        prefix<- "Z://" 
      } #NFWF Server ran out of space so some temporarily on Scamp2 D Drive
    if(!file.exists(paste0(prefix,mb_df$Server_Location[i]))){stop("Error: File does not exist")}
    
    r<- rast(paste0(prefix,mb_df$Server_Location[i]))
    
    if(st_is_empty(mb_df[i,])){
      shp<- (r < Inf) %>% 
      as.polygons(dissolve=TRUE, values=FALSE, na.rm=TRUE) %>%
      fillHoles() %>% 
      st_as_sf() %>%
      st_transform("EPSG:4326")
      mb_df$geometry[i]<- st_geometry(shp)
    } # Only generate footprints if it doesn't exist
    
    proj_string<- as.character(crs(r, proj=TRUE))
    Projection<- suppressWarnings(str_extract(proj_string, regex("(?<=proj=)[[:alpha:]]+", perl=TRUE)))
    Zone<-suppressWarnings(str_extract(proj_string, regex("(?<=zone=)[[:digit:]]+", perl=TRUE)))
    Ellipsoid<-suppressWarnings(str_extract(proj_string, regex("(?<=datum=)[[:alnum:]]+", perl=TRUE)))
    if(is.na(Ellipsoid)){
      Ellipsoid<- suppressWarnings(str_extract(proj_string, regex("(?<=ellps=)[[:alnum:]]+", perl=TRUE)))
    }
    if(is.na(Ellipsoid)){
      Ellipsoid<- paste(unlist(str_extract_all(proj_string, "\\+((rf)|(f)|(a)|(b))=[^ ]+")), collapse = " ")
    }
    if(!is.na(Zone)){
      Projection<- paste(Projection, Zone)
    }
    Units<- suppressWarnings(str_extract(proj_string, regex("(?<=units=)[[:alpha:]]+", perl=TRUE)))
    
    mb_df$Filename[i]<- basename(mb_df$Server_Location[i])
    mb_df$Resolution[i]<- xres(r)
    mb_df$Units[i]<- Units
    mb_df$Projection_Description[i]<- as.character(crs(r, describe=TRUE, parse=TRUE)[1])
    mb_df$Projection[i]<- Projection
    mb_df$Ellipsoid[i]<- Ellipsoid
  }}

mb_df<- mb_df %>% select(-Old_Server_Location) # drop column

write_csv(st_drop_geometry(mb_df), "Multibeam_metadata.csv", append=FALSE)
write_xlsx(st_drop_geometry(mb_df), "Multibeam_metadata.xlsx")
st_write(mb_df, "Multibeam_metadata.gpkg", append = FALSE)
st_write(mb_df, "Multibeam_metadata.shp", append = FALSE)

write_csv(st_drop_geometry(mb_df), "Z://2_Projects/GIS/Multibeam_Metadata_And_Footprints_AI/Multibeam_metadata.csv", append=FALSE)
write_xlsx(st_drop_geometry(mb_df), "Z://2_Projects/GIS/Multibeam_Metadata_And_Footprints_AI/Multibeam_metadata.xlsx")
st_write(mb_df, "Z://2_Projects/GIS/Multibeam_Metadata_And_Footprints_AI/Multibeam_metadata.gpkg", append = FALSE)
st_write(mb_df, "Z://2_Projects/GIS/Multibeam_Metadata_And_Footprints_AI/Multibeam_metadata.shp", append = FALSE)
```

```{r}
options(knitr.kable.NA = '')
knitr::kable(st_drop_geometry(mb_df), format = "pipe", digits = 5, align = "c")
```

